---
title: "Application of Spatial and Spatio-temporal Point Patterns Analysis to discover the geographical distribution of Armed Conflict in Myanmar"
author: "Santhya Selvan"
date: "September 4, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## 1.0 Overview

Myanmar has been going through civil wars and internal conflicts for decades now. These conflicts have mainly been due to ethnic tensions and general unhappiness in the governing party. Conflicts range from battles, riots, violence against civilians, strategic developments and more. In this exercise, I will be analysing the distribution of conflicts in Myanmar.

To do this, I will mainly be performing quarterly KDE analysis on 4 main types of conflicts: Battles, Explosions, Violence against Civillians and Strategic Developments through the years of 2021-2024. I will also be performing the second order spatial point pattern process as well as the spatio temporal point process analysis for these segments. Let's get started.

#### 1.1 Loading R packages

These are the packages that I will be using in this exercise.

**sf**: Used in spatial data wrangling

**tidyverse**: Used in data wrangling for non-spatial data

**raster**: For reading, writing and manipulating raster data

**tmap**: For functions relating to mapping point patterns

**spatstat**: Provides functions for spatial point process analysis

**sparr**: Provides functions for spatio-temporal point process analysis

```{r}
pacman::p_load(sf, raster, tidyverse, spatstat, sparr, tmap)
```

## 2.0 Data Preparation

In this section, I will be retrieving the required data and performing the necessary data wrangling methods in order to transform the data into a form that is best suitable for our analytic needs for this exercise. The datasets used in this exercise are:

-   `Myanmar_All_2021-24.csv`: A csv file downloaded from ACLED, which contains all the details of the internal conflicts in Myanmar from 2021-2024. Note that the file name has been renamed for ease of use.

-   `MBoundary`: A shapefile downloaded from Myanmar Information Management Unit, MIMU. This shapefile maps out the boundary of Myanmar. I have also renamed this file for ease of use.

#### 2.1 Importing the Data

Firstly, I will be importing the base map of Myanmar in shapefile format. I will also transform the crs to that of Myanmar.

```{r}
boundary <- st_read(dsn="data/raw",
                    layer = "MBoundary") %>% 
  st_transform(crs=32646)

```

```{r}
st_crs(boundary)
```

Next, I will also be Importing the regional boundaries of Myanmar.

```{r}
regions_sf <- st_read(dsn="data/raw",
                      layer="regions") %>% 
  st_transform(crs = 32646)
```

Lastly, I will be importing and reading the csv file containing the data about the conflicts into a dataframe. I will once again ensure that the crs of the data is transformed to that of Myanmar's.

```{r}

all_sf <- read_csv('data/raw/Myanmar_All_2021-24.csv')%>% 
  st_as_sf(coords = c('longitude', 'latitude'),
           crs=4326) %>% 
  st_transform(crs=32646)

st_crs(all_sf)
```

#### 2.2 Preparing the Data

Before we can start deriving our KDE layers, we will make sure our data is in a suitable form and is processed in a way that we need it in. I will start off by converting the date column in the aspatial data to a format that can be handled during computations.

```{r}

all_sf$event_date <- as.Date(all_sf$event_date, format = "%d %B %Y")
```

Before we can filter out the data that we need, I will be visualising the data according to the event types to observe for any patterns or similarity.

```{r}

tm_shape(boundary) + tm_polygons() + tm_shape(regions_sf) + tm_polygons() + tm_shape(all_sf) + tm_dots() + tm_facets(by='event_type', free.coords = FALSE, drop.units= TRUE)
```

From the patterns observed above, we see that Battles, Strategic Developments, Explosions and Violence against civilians all have similar distribution of conflicts across the space and that Riots and Protests do not have that many spatial points even over the aggregated 4 year dataset. As such, I will be focusing my analysis on these four main categories of conflicts: **Battles, Strategic Developments, Violence against Civilians and Explosions.**

I will filter out the data according to these event types and save them in separate sf objects.

```{r}
#| eval: false
battles_sf <- all_sf %>%  filter(event_type == 'Battles')
battles_sf

exp_sf <- all_sf %>%  filter(event_type == "Explosions/Remote violence")

civViolence_sf <- all_sf %>%  filter(event_type == "Violence against civilians")

strat_dev_sf <- all_sf %>%  filter(event_type == 'Strategic developments')


```

Finally, I will further split the datasets into quarterly periods. For ease of use, I will write these sf objects into rds files and save them in a compiled folder so I won't mix them up with the unprocessed data.

::: panel-tabset
## Battles

```{r}
#| eval: false
q1_21_battles_sf <- battles_sf %>% 
  filter(year == '2021' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Battles/2021/q1.rds")

q2_21_battles_sf <-battles_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Battles/2021/q2.rds")

q3_21_battles_sf <-battles_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Battles/2021/q3.rds")

q4_21_battles_sf <-battles_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Battles/2021/q4.rds")

#2022

q1_22_battles_sf <- battles_sf %>% 
  filter(year == '2022' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Battles/2022/q1.rds")

q2_22_battles_sf <-battles_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Battles/2022/q2.rds")

q3_22_battles_sf <-battles_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Battles/2022/q3.rds")

q4_22_battles_sf <-battles_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Battles/2022/q4.rds")

#2023

q1_23_battles_sf <- battles_sf %>% 
  filter(year == '2023' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Battles/2023/q1.rds")

q2_23_battles_sf <-battles_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Battles/2023/q2.rds")

q3_23_battles_sf <-battles_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Battles/2023/q3.rds")

q4_23_battles_sf <-battles_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Battles/2023/q4.rds")

#2024

q1_24_battles_sf <- battles_sf %>% 
  filter(year == '2024' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Battles/2024/q1.rds")

q2_24_battles_sf <-battles_sf %>% 
  filter(year == '2024') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Battles/2024/q2.rds")

```

## Explosions

```{r}
#| eval: false

q1_21_exp_sf <- exp_sf %>% 
  filter(year == '2021' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Explosions/2021/q1.rds")

q2_21_exp_sf <-exp_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Explosions/2021/q2.rds")

q3_21_exp_sf <-exp_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Explosions/2021/q3.rds")

q4_21_exp_sf <-exp_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Explosions/2021/q4.rds")

#2022

q1_22_exp_sf <- exp_sf %>% 
  filter(year == '2022' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Explosions/2022/q1.rds")

q2_22_exp_sf <-exp_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Explosions/2022/q2.rds")

q3_22_exp_sf <-exp_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Explosions/2022/q3.rds")

q4_22_exp_sf <-exp_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Explosions/2022/q4.rds")

#2023

q1_23_exp_sf <-exp_sf %>% 
  filter(year == '2023' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Explosions/2023/q1.rds")

q2_23_exp_sf <-exp_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Explosions/2023/q2.rds")

q3_23_exp_sf <-exp_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Explosions/2023/q3.rds")

q4_23_exp_sf <-battles_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Explosions/2023/q4.rds")

#2024

q1_24_exp_sf <- exp_sf %>% 
  filter(year == '2024' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Explosions/2024/q1.rds")

q2_24_exp_sf <-exp_sf %>% 
  filter(year == '2024') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Explosions/2024/q2.rds")
```

## Violence against civillians

```{r}
#| eval: false
q1_21_civViolence_sf <- civViolence_sf %>% 
  filter(year == '2021' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Violence/2021/q1.rds")

q2_21_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Violence/2021/q2.rds")

q3_21_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Violence/2021/q3.rds")

q4_21_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Violence/2021/q4.rds")

#2022

q1_22_civViolence_sf <- civViolence_sf %>% 
  filter(year == '2022' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Violence/2022/q1.rds")

q2_22_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Violence/2022/q2.rds")

q3_22_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Violence/2022/q3.rds")

q4_22_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Violence/2022/q4.rds")

#2023

q1_23_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2023' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Violence/2023/q1.rds")

q2_23_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Violence/2023/q2.rds")

q3_23_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Violence/2023/q3.rds")

q4_23_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Violence/2023/q4.rds")

#2024

q1_24_civViolence_sf <- civViolence_sf %>% 
  filter(year == '2024' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Violence/2024/q1.rds")

q2_24_civViolence_sf <-civViolence_sf %>% 
  filter(year == '2024') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Violence/2024/q2.rds")
```

## Strategic Developments

```{r}
#| eval: false
q1_21_strat_dev_sf <- strat_dev_sf %>% 
  filter(year == '2021' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Strat_Dev/2021/q1.rds")

q2_21_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Strat_Dev/2021/q2.rds")

q3_21_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Strat_Dev/2021/q3.rds")

q4_21_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2021') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Strat_Dev/2021/q4.rds")

#2022

q1_22_strat_dev_sf <- strat_dev_sf %>% 
  filter(year == '2022' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Strat_Dev/2022/q1.rds")

q2_22_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Strat_Dev/2022/q2.rds")

q3_22_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Strat_Dev/2022/q3.rds")

q4_22_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2022') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Strat_Dev/2022/q4.rds")

#2023

q1_23_strat_dev_sf <-civViolence_sf %>% 
  filter(year == '2023' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Strat_Dev/2023/q1.rds")

q2_23_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Strat_Dev/2023/q2.rds")

q3_23_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 3)%>%  
  write_rds("data/rds/Strat_Dev/2023/q3.rds")

q4_23_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2023') %>% 
  filter(quarter(event_date) == 4)%>%  
  write_rds("data/rds/Strat_Dev/2023/q4.rds")

#2024

q1_24_strat_dev_sf <- strat_dev_sf %>% 
  filter(year == '2024' ) %>% 
  filter(quarter(event_date) == 1) %>%  
  write_rds("data/rds/Strat_Dev/2024/q1.rds")

q2_24_strat_dev_sf <-strat_dev_sf %>% 
  filter(year == '2024') %>% 
  filter(quarter(event_date) == 2)%>%  
  write_rds("data/rds/Strat_Dev/2024/q2.rds")
```
:::

```{r}
#| echo: false
#Battles
q1_21_battles_sf<-read_rds("data/rds/Battles/2021/q1.rds")
q2_21_battles_sf<-read_rds("data/rds/Battles/2021/q2.rds")
q3_21_battles_sf<-read_rds("data/rds/Battles/2021/q3.rds")
q4_21_battles_sf<-read_rds("data/rds/Battles/2021/q4.rds")

q1_22_battles_sf<-read_rds("data/rds/Battles/2022/q1.rds")
q2_22_battles_sf<-read_rds("data/rds/Battles/2022/q2.rds")
q3_22_battles_sf<-read_rds("data/rds/Battles/2022/q3.rds")
q4_22_battles_sf<-read_rds("data/rds/Battles/2022/q4.rds")

q1_23_battles_sf<-read_rds("data/rds/Battles/2023/q1.rds")
q2_23_battles_sf<-read_rds("data/rds/Battles/2023/q2.rds")
q3_23_battles_sf<-read_rds("data/rds/Battles/2023/q3.rds")
q4_23_battles_sf<-read_rds("data/rds/Battles/2023/q4.rds")

q1_24_battles_sf<-read_rds("data/rds/Battles/2024/q1.rds")
q2_24_battles_sf<-read_rds("data/rds/Battles/2024/q2.rds")

#Explosions

q1_21_exp_sf<-read_rds("data/rds/Explosions/2021/q1.rds")
q2_21_exp_sf<-read_rds("data/rds/Explosions/2021/q2.rds")
q3_21_exp_sf<-read_rds("data/rds/Explosions/2021/q3.rds")
q4_21_exp_sf<-read_rds("data/rds/Explosions/2021/q4.rds")

q1_22_exp_sf<-read_rds("data/rds/Explosions/2022/q1.rds")
q2_22_exp_sf<-read_rds("data/rds/Explosions/2022/q2.rds")
q3_22_exp_sf<-read_rds("data/rds/Explosions/2022/q3.rds")
q4_22_exp_sf<-read_rds("data/rds/Explosions/2022/q4.rds")

q1_23_exp_sf<-read_rds("data/rds/Explosions/2023/q1.rds")
q2_23_exp_sf<-read_rds("data/rds/Explosions/2023/q2.rds")
q3_23_exp_sf<-read_rds("data/rds/Explosions/2023/q3.rds")
q4_23_exp_sf<-read_rds("data/rds/Explosions/2023/q4.rds")

q1_24_exp_sf<-read_rds("data/rds/Explosions/2024/q1.rds")
q2_24_exp_sf<-read_rds("data/rds/Explosions/2024/q2.rds")


#Strat Dev

q1_21_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2021/q1.rds")
q2_21_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2021/q2.rds")
q3_21_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2021/q3.rds")
q4_21_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2021/q4.rds")

q1_22_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2022/q1.rds")
q2_22_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2022/q2.rds")
q3_22_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2022/q3.rds")
q4_22_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2022/q4.rds")

q1_23_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2023/q1.rds")
q2_23_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2023/q2.rds")
q3_23_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2023/q3.rds")
q4_23_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2023/q4.rds")

q1_24_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2024/q1.rds")
q2_24_strat_dev_sf<-read_rds("data/rds/Strat_Dev/2024/q2.rds")

#Violence

q1_21_civViolence_sf<-read_rds("data/rds/Violence/2021/q1.rds")
q2_21_civViolence_sf<-read_rds("data/rds/Violence/2021/q2.rds")
q3_21_civViolence_sf<-read_rds("data/rds/Violence/2021/q3.rds")
q4_21_civViolence_sf<-read_rds("data/rds/Violence/2021/q4.rds")

q1_22_civViolence_sf<-read_rds("data/rds/Violence/2022/q1.rds")
q2_22_civViolence_sf<-read_rds("data/rds/Violence/2022/q2.rds")
q3_22_civViolence_sf<-read_rds("data/rds/Violence/2022/q3.rds")
q4_22_civViolence_sf<-read_rds("data/rds/Violence/2022/q4.rds")

q1_23_civViolence_sf<-read_rds("data/rds/Violence/2023/q1.rds")
q2_23_civViolence_sf<-read_rds("data/rds/Violence/2023/q2.rds")
q3_23_civViolence_sf<-read_rds("data/rds/Violence/2023/q3.rds")
q4_23_civViolence_sf<-read_rds("data/rds/Violence/2023/q4.rds")

q1_24_civViolence_sf<-read_rds("data/rds/Violence/2024/q1.rds")
q2_24_civViolence_sf<-read_rds("data/rds/Violence/2024/q2.rds")
```

#### 2.2 Data Wrangling

##### 2.2.1 Converting the sf objects to ppp objects

To derive our KDE layers, we would need the sf objects to be converted to ppp objects. We will do this using the *as.ppp()* function.

::: panel-tabset
## Battles

```{r}
battlesQ121_ppp <- as.ppp(st_coordinates(q1_21_battles_sf), st_bbox(q1_21_battles_sf))
battlesQ221_ppp <- as.ppp(st_coordinates(q2_21_battles_sf), st_bbox(q2_21_battles_sf))
battlesQ321_ppp<- as.ppp(st_coordinates(q3_21_battles_sf), st_bbox(q3_21_battles_sf))
battlesQ421_ppp<-as.ppp(st_coordinates(q4_21_battles_sf), st_bbox(q4_21_battles_sf))

#2022

battlesQ122_ppp <- as.ppp(st_coordinates(q1_22_battles_sf), st_bbox(q1_22_battles_sf))
battlesQ222_ppp <- as.ppp(st_coordinates(q2_22_battles_sf), st_bbox(q2_22_battles_sf))
battlesQ322_ppp<- as.ppp(st_coordinates(q3_22_battles_sf), st_bbox(q3_22_battles_sf))
battlesQ422_ppp<-as.ppp(st_coordinates(q4_22_battles_sf), st_bbox(q4_22_battles_sf))

#2023
battlesQ123_ppp <- as.ppp(st_coordinates(q1_23_battles_sf), st_bbox(q1_23_battles_sf))
battlesQ223_ppp <- as.ppp(st_coordinates(q2_23_battles_sf), st_bbox(q2_23_battles_sf))
battlesQ323_ppp<- as.ppp(st_coordinates(q3_23_battles_sf), st_bbox(q3_23_battles_sf))
battlesQ423_ppp<-as.ppp(st_coordinates(q4_23_battles_sf), st_bbox(q4_23_battles_sf))
  
#2024
battlesQ124_ppp <- as.ppp(st_coordinates(q1_24_battles_sf), st_bbox(q1_24_battles_sf))
battlesQ224_ppp <- as.ppp(st_coordinates(q2_24_battles_sf), st_bbox(q2_24_battles_sf))
  
```

## Explosions

```{r}
expQ121_ppp <- as.ppp(st_coordinates(q1_21_exp_sf), st_bbox(q1_21_exp_sf))
expQ221_ppp <- as.ppp(st_coordinates(q2_21_exp_sf), st_bbox(q2_21_exp_sf))
expQ321_ppp<- as.ppp(st_coordinates(q3_21_exp_sf), st_bbox(q3_21_exp_sf))
expQ421_ppp<-as.ppp(st_coordinates(q4_21_exp_sf), st_bbox(q4_21_exp_sf))

#2022

expQ122_ppp <- as.ppp(st_coordinates(q1_22_exp_sf), st_bbox(q1_22_exp_sf))
expQ222_ppp <- as.ppp(st_coordinates(q2_22_exp_sf), st_bbox(q2_22_exp_sf))
expQ322_ppp<- as.ppp(st_coordinates(q3_22_exp_sf), st_bbox(q3_22_exp_sf))
expQ422_ppp<-as.ppp(st_coordinates(q4_22_exp_sf), st_bbox(q4_22_exp_sf))

#2023
expQ123_ppp <- as.ppp(st_coordinates(q1_23_exp_sf), st_bbox(q1_23_exp_sf))
expQ223_ppp <- as.ppp(st_coordinates(q2_23_exp_sf), st_bbox(q2_23_exp_sf))
expQ323_ppp<- as.ppp(st_coordinates(q3_23_exp_sf), st_bbox(q3_23_exp_sf))
expQ423_ppp<-as.ppp(st_coordinates(q4_23_exp_sf), st_bbox(q4_23_exp_sf))
  
#2024
expQ124_ppp <- as.ppp(st_coordinates(q1_24_exp_sf), st_bbox(q1_24_exp_sf))
expQ224_ppp <- as.ppp(st_coordinates(q2_24_exp_sf), st_bbox(q2_24_exp_sf))
```

## Strategic Developments

```{r}
strat_devQ121_ppp <- as.ppp(st_coordinates(q1_21_strat_dev_sf), st_bbox(q1_21_strat_dev_sf))
strat_devQ221_ppp <- as.ppp(st_coordinates(q2_21_strat_dev_sf), st_bbox(q2_21_strat_dev_sf))
strat_devQ321_ppp<- as.ppp(st_coordinates(q3_21_strat_dev_sf), st_bbox(q3_21_strat_dev_sf))
strat_devQ421_ppp<-as.ppp(st_coordinates(q4_21_strat_dev_sf), st_bbox(q4_21_strat_dev_sf))

#2022
strat_devQ122_ppp <- as.ppp(st_coordinates(q1_22_strat_dev_sf), st_bbox(q1_22_strat_dev_sf))
strat_devQ222_ppp <- as.ppp(st_coordinates(q2_22_strat_dev_sf), st_bbox(q2_22_strat_dev_sf))
strat_devQ322_ppp<- as.ppp(st_coordinates(q3_22_strat_dev_sf), st_bbox(q3_22_strat_dev_sf))
strat_devQ422_ppp<-as.ppp(st_coordinates(q4_22_strat_dev_sf), st_bbox(q4_22_strat_dev_sf))

#2023
strat_devQ123_ppp <- as.ppp(st_coordinates(q1_23_strat_dev_sf), st_bbox(q1_23_strat_dev_sf))
strat_devQ223_ppp <- as.ppp(st_coordinates(q2_23_strat_dev_sf), st_bbox(q2_23_strat_dev_sf))
strat_devQ323_ppp<- as.ppp(st_coordinates(q3_23_strat_dev_sf), st_bbox(q3_23_strat_dev_sf))
strat_devQ423_ppp<-as.ppp(st_coordinates(q4_23_strat_dev_sf), st_bbox(q4_23_strat_dev_sf))
  
#2024
strat_devQ124_ppp <- as.ppp(st_coordinates(q1_24_strat_dev_sf), st_bbox(q1_24_strat_dev_sf))
strat_devQ224_ppp <- as.ppp(st_coordinates(q2_24_strat_dev_sf), st_bbox(q2_24_strat_dev_sf))
```

## Violence Against Civilians

```{r}
civViolenceQ121_ppp <- as.ppp(st_coordinates(q1_21_civViolence_sf), st_bbox(q1_21_civViolence_sf))
civViolenceQ221_ppp <- as.ppp(st_coordinates(q2_21_civViolence_sf), st_bbox(q2_21_civViolence_sf))
civViolenceQ321_ppp<- as.ppp(st_coordinates(q3_21_civViolence_sf), st_bbox(q3_21_civViolence_sf))
civViolenceQ421_ppp<-as.ppp(st_coordinates(q4_21_civViolence_sf), st_bbox(q4_21_civViolence_sf))

#2022

civViolenceQ122_ppp <- as.ppp(st_coordinates(q1_22_civViolence_sf), st_bbox(q1_22_civViolence_sf))
civViolenceQ222_ppp <- as.ppp(st_coordinates(q2_22_civViolence_sf), st_bbox(q2_22_civViolence_sf))
civViolenceQ322_ppp<- as.ppp(st_coordinates(q3_22_civViolence_sf), st_bbox(q3_22_civViolence_sf))
civViolenceQ422_ppp<-as.ppp(st_coordinates(q4_22_civViolence_sf), st_bbox(q4_22_civViolence_sf))

#2023
civViolenceQ123_ppp <- as.ppp(st_coordinates(q1_23_civViolence_sf), st_bbox(q1_23_civViolence_sf))
civViolenceQ223_ppp <- as.ppp(st_coordinates(q2_23_civViolence_sf), st_bbox(q2_23_civViolence_sf))
civViolenceQ323_ppp<- as.ppp(st_coordinates(q3_23_civViolence_sf), st_bbox(q3_23_civViolence_sf))
civViolenceQ423_ppp<-as.ppp(st_coordinates(q4_23_civViolence_sf), st_bbox(q4_23_civViolence_sf))
  
#2024
civViolenceQ124_ppp <- as.ppp(st_coordinates(q1_24_civViolence_sf), st_bbox(q1_24_civViolence_sf))
civViolenceQ224_ppp <- as.ppp(st_coordinates(q2_24_civViolence_sf), st_bbox(q2_24_civViolence_sf))
```
:::

#### 2.2.2 Converting the boundary to an owin object

```{r}
boundary_owin <- as.owin(boundary)
```

Once we have our ppp object and owin object, we can start combining them to one object that can be passed through the *density()* function.

::: panel-tabset
## Battles

```{r}
battlesQ121_ppp<- rescale(battlesQ121_ppp, 1000, "km") 
battlesQ121_ppp_com<- battlesQ121_ppp[boundary_owin]

battlesQ221_ppp<- rescale(battlesQ221_ppp, 1000, "km") 
battlesQ221_ppp_com<- battlesQ221_ppp[boundary_owin]

battlesQ321_ppp<- rescale(battlesQ321_ppp, 1000, "km") 
battlesQ3121_ppp_com<- battlesQ321_ppp[boundary_owin]

battlesQ421_ppp<- rescale(battlesQ421_ppp, 1000, "km") 
battlesQ421_ppp_com<- battlesQ421_ppp[boundary_owin]

#2022
battlesQ122_ppp<- rescale(battlesQ122_ppp, 1000, "km") 
battlesQ122_ppp_com<- battlesQ122_ppp[boundary_owin]

battlesQ222_ppp<- rescale(battlesQ222_ppp, 1000, "km") 
battlesQ222_ppp_com<- battlesQ222_ppp[boundary_owin]

battlesQ322_ppp<- rescale(battlesQ322_ppp, 1000, "km") 
battlesQ3122_ppp_com<- battlesQ322_ppp[boundary_owin]

battlesQ422_ppp<- rescale(battlesQ422_ppp, 1000, "km") 
battlesQ422_ppp_com<- battlesQ422_ppp[boundary_owin]
                                                                                #2023
                                      
battlesQ123_ppp<- rescale(battlesQ123_ppp, 1000, "km") 
battlesQ123_ppp_com<- battlesQ123_ppp[boundary_owin]

battlesQ223_ppp<- rescale(battlesQ223_ppp, 1000, "km") 
battlesQ223_ppp_com<- battlesQ223_ppp[boundary_owin]

battlesQ323_ppp<- rescale(battlesQ323_ppp, 1000, "km") 
battlesQ3123_ppp_com<- battlesQ323_ppp[boundary_owin]

battlesQ423_ppp<- rescale(battlesQ423_ppp, 1000, "km") 
battlesQ423_ppp_com<- battlesQ423_ppp[boundary_owin]

#2024

battlesQ124_ppp<- rescale(battlesQ124_ppp, 1000, "km") 
battlesQ124_ppp_com<- battlesQ124_ppp[boundary_owin]

battlesQ224_ppp<- rescale(battlesQ224_ppp, 1000, "km") 
battlesQ224_ppp_com<- battlesQ224_ppp[boundary_owin]
```

## Explosions

```{r}
expQ121_ppp<- rescale(expQ121_ppp, 1000, "km") 
expQ121_ppp_com<- expQ121_ppp[boundary_owin]

expQ221_ppp<- rescale(expQ221_ppp, 1000, "km") 
expQ221_ppp_com<- expQ221_ppp[boundary_owin]

expQ321_ppp<- rescale(expQ321_ppp, 1000, "km") 
expQ3121_ppp_com<- expQ321_ppp[boundary_owin]

expQ421_ppp<- rescale(expQ421_ppp, 1000, "km") 
expQ421_ppp_com<- expQ421_ppp[boundary_owin]

#2022
expQ122_ppp<- rescale(expQ122_ppp, 1000, "km") 
expQ122_ppp_com<- expQ122_ppp[boundary_owin]

expQ222_ppp<- rescale(expQ222_ppp, 1000, "km") 
expQ222_ppp_com<- expQ222_ppp[boundary_owin]

expQ322_ppp<- rescale(expQ322_ppp, 1000, "km") 
expQ3122_ppp_com<- expQ322_ppp[boundary_owin]

expQ422_ppp<- rescale(expQ422_ppp, 1000, "km") 
expQ422_ppp_com<- expQ422_ppp[boundary_owin]
                                      
#2023
                                      
expQ123_ppp<- rescale(expQ123_ppp, 1000, "km") 
expQ123_ppp_com<- expQ123_ppp[boundary_owin]

expQ223_ppp<- rescale(expQ223_ppp, 1000, "km") 
expQ223_ppp_com<- expQ223_ppp[boundary_owin]

expQ323_ppp<- rescale(expQ323_ppp, 1000, "km") 
expQ3123_ppp_com<- expQ323_ppp[boundary_owin]

expQ423_ppp<- rescale(expQ423_ppp, 1000, "km") 
expQ423_ppp_com<- expQ423_ppp[boundary_owin]

#2024

expQ124_ppp<- rescale(expQ124_ppp, 1000, "km") 
expQ124_ppp_com<- expQ124_ppp[boundary_owin]
```

## Strategic Developments

```{r}
strat_devQ121_ppp<- rescale(strat_devQ121_ppp, 1000, "km") 
strat_devQ121_ppp_com<- strat_devQ121_ppp[boundary_owin]

strat_devQ221_ppp<- rescale(strat_devQ221_ppp, 1000, "km") 
strat_devQ221_ppp_com<- strat_devQ221_ppp[boundary_owin]

strat_devQ321_ppp<- rescale(strat_devQ321_ppp, 1000, "km") 
strat_devQ3121_ppp_com<- strat_devQ321_ppp[boundary_owin]

strat_devQ421_ppp<- rescale(strat_devQ421_ppp, 1000, "km") 
strat_devQ421_ppp_com<- strat_devQ421_ppp[boundary_owin]

#2022
strat_devQ122_ppp<- rescale(strat_devQ122_ppp, 1000, "km") 
strat_devQ122_ppp_com<- strat_devQ122_ppp[boundary_owin]

strat_devQ222_ppp<- rescale(strat_devQ222_ppp, 1000, "km") 
strat_devQ222_ppp_com<- strat_devQ222_ppp[boundary_owin]

strat_devQ322_ppp<- rescale(strat_devQ322_ppp, 1000, "km") 
strat_devQ3122_ppp_com<- strat_devQ322_ppp[boundary_owin]

strat_devQ422_ppp<- rescale(strat_devQ422_ppp, 1000, "km") 
strat_devQ422_ppp_com<- strat_devQ422_ppp[boundary_owin]

#2023
                                      
strat_devQ123_ppp<- rescale(strat_devQ123_ppp, 1000, "km") 
strat_devQ123_ppp_com<- strat_devQ123_ppp[boundary_owin]

strat_devQ223_ppp<- rescale(strat_devQ223_ppp, 1000, "km") 
strat_devQ223_ppp_com<- strat_devQ223_ppp[boundary_owin]

strat_devQ323_ppp<- rescale(strat_devQ323_ppp, 1000, "km") 
strat_devQ3123_ppp_com<- strat_devQ323_ppp[boundary_owin]

strat_devQ423_ppp<- rescale(strat_devQ423_ppp, 1000, "km") 
strat_devQ423_ppp_com<- strat_devQ423_ppp[boundary_owin]

#2024

strat_devQ124_ppp<- rescale(strat_devQ124_ppp, 1000, "km") 
strat_devQ124_ppp_com<- strat_devQ124_ppp[boundary_owin]
```

## Violence Against Civilians

```{r}
civViolenceQ121_ppp<- rescale(civViolenceQ121_ppp, 1000, "km") 
civViolenceQ121_ppp_com<- civViolenceQ121_ppp[boundary_owin]

civViolenceQ221_ppp<- rescale(civViolenceQ221_ppp, 1000, "km") 
civViolenceQ221_ppp_com<- civViolenceQ221_ppp[boundary_owin]

civViolenceQ321_ppp<- rescale(civViolenceQ321_ppp, 1000, "km") 
civViolenceQ3121_ppp_com<- civViolenceQ321_ppp[boundary_owin]

civViolenceQ421_ppp<- rescale(civViolenceQ421_ppp, 1000, "km") 
civViolenceQ421_ppp_com<- civViolenceQ421_ppp[boundary_owin]

#2022
civViolenceQ122_ppp<- rescale(civViolenceQ122_ppp, 1000, "km") 
civViolenceQ122_ppp_com<- civViolenceQ122_ppp[boundary_owin]

civViolenceQ222_ppp<- rescale(civViolenceQ222_ppp, 1000, "km") 
civViolenceQ222_ppp_com<- civViolenceQ222_ppp[boundary_owin]

civViolenceQ322_ppp<- rescale(civViolenceQ322_ppp, 1000, "km") 
civViolenceQ3122_ppp_com<- civViolenceQ322_ppp[boundary_owin]

civViolenceQ422_ppp<- rescale(civViolenceQ422_ppp, 1000, "km") 
civViolenceQ422_ppp_com<- civViolenceQ422_ppp[boundary_owin]

#2023
                                      
civViolenceQ123_ppp<- rescale(civViolenceQ123_ppp, 1000, "km") 
civViolenceQ123_ppp_com<- civViolenceQ123_ppp[boundary_owin]

civViolenceQ223_ppp<- rescale(civViolenceQ223_ppp, 1000, "km") 
civViolenceQ223_ppp_com<- civViolenceQ223_ppp[boundary_owin]

civViolenceQ323_ppp<- rescale(civViolenceQ323_ppp, 1000, "km") 
civViolenceQ3123_ppp_com<- civViolenceQ323_ppp[boundary_owin]

civViolenceQ423_ppp<- rescale(civViolenceQ423_ppp, 1000, "km") 
civViolenceQ423_ppp_com<- civViolenceQ423_ppp[boundary_owin]

#2024

civViolenceQ124_ppp<- rescale(civViolenceQ124_ppp, 1000, "km") 
civViolenceQ124_ppp_com<- civViolenceQ124_ppp[boundary_owin]
```
:::

## 3.0 Computing the quarterly KDE Layers

```{r}
#battles_kde <- density(battles_ppp_com, sigma=bw.CvL(battles_ppp_com), edge=TRUE, kernel="gaussian") 

#plot(battles_kde)

```
