---
title: "Application of Spatial and Spatio-temporal Point Patterns Analysis to discover the geographical distribution of Armed Conflict in Myanmar"
author: "Santhya Selvan"
date: "September 4, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## 1.0 Overview

......

#### 1.1 Loading R packages

These are the packages that I will be using in this exercise.

**sf**: Used in spatial data wrangling

**tidyverse**: Used in data wrangling for non-spatial data

**raster**: For reading, writing and manipulating raster data

**tmap**: For functions relating to mapping point patterns

**spatstat**: Provides functions for spatial point process analysis

**sparr**: Provides functions for spatio-temporal point process analysis

```{r}
pacman::p_load(sf, raster, tidyverse, spatstat, sparr, tmap)
```

## 2.0 Data Preparation

In this section, I will be retrieving the required data and performing the necessary data wrangling methods in order to transform the data into a form that is best suitable for our analytic needs for this exercise. The datasets used in this exercise are:

-   `Myanmar_All_2021-24.csv`: A csv file downloaded from ACLED, which contains all the details of the internal conflicts in Myanmar from 2021-2024. Note that the file name has been renamed for ease of use.

-   `MBoundary`: A shapefile downloaded from Myanmar Information Management Unit, MIMU. This shapefile maps out the boundary of Myanmar. I have also renamed this file for ease of use.

#### 2.1 Importing the Data

importing the base map of myanmar/ geospatial data. transform the crs to that of Myanmar.

```{r}
boundary <- st_read(dsn="data/raw",
                    layer = "MBoundary") %>% 
  st_transform(crs=32646)

```

```{r}
st_crs(boundary)
```

Importing the regional boundaries of Myanmar

```{r}
regions_sf <- st_read(dsn="data/raw",
                      layer="regions") %>% 
  st_transform(crs = 32646)
```

importing and reading the csv file containing the data about the conflicts into a dataframe. transform the crs to that of Myanmars

```{r}
all_sf <- read_csv('data/raw/Myanmar_All_2021-24.csv')%>% 
  st_as_sf(coords = c('longitude', 'latitude'),
           crs=4326) %>% 
  st_transform(crs=32646)

st_crs(all_sf)
```

converting the date format to be more usable.

```{r}
all_sf$event_date <- as.Date(all_sf$event_date, format = "%d %B %Y")
```

filter out the event types we will be focusing on in the exercise and save them into different dataframes.

```{r}
battles_sf <- all_sf %>%  filter(event_type == 'Battles')
battles_sf

exp_sf <- all_sf %>%  filter(event_type == "Explosions/Remote violence")

civViolence_sf <- all_sf %>%  filter(event_type == "Violence against civilians")

riots_sf <- all_sf %>%  filter(event_type=='Riots')

strat_dev_sf <- all_sf %>%  filter(event_type == 'Strategic Developments')


```

mapping out the spatial data and the spatial points by the event type to observe for any areas of interest.

```{r}
tm_shape(boundary) + tm_polygons() + tm_shape(regions_sf) + tm_polygons() + tm_shape(all_sf) + tm_dots() + tm_facets(by='event_type', free.coords = FALSE, drop.units= TRUE)
```

split the dataset into quarterly

::: panel-tabset
## Battles

## Explosions

## Violence against civillians

## Strategic Developments
:::

```{r}
q1_21_battles_sf <- battles_sf %>% 
  filter(year == '2021' ) %>% 
  filter(quarter(event_date) == 1)

```

```{r}
tm_shape(boundary) + tm_polygons() + tm_shape(regions_sf) + tm_polygons() + tm_shape(q1_21_battles_sf) + tm_dots()
```

#### 2.2 Data Wrangling

##### 2.2.1 Converting the sf objects to ppp objects using

```{r}
battles_ppp <- as.ppp(st_coordinates(q1_21_battles_sf), st_bbox(q1_21_battles_sf))
```

##### 2.2.2 Converting the boundary to an owin object

```{r}
boundary_owin <- as.owin(boundary)
```

combining the ppp object to the owin object

```{r}
battles_ppp_com <- battles_ppp[boundary_owin]
```

```{r}
plot(battles_ppp_com)
```

test plotting the kde layer for battles for q1 of 2021

```{r}
battles_ppp_com.km <- rescale.ppp(battles_ppp_com, 1000, 'km')
```

## 3.0 Computing the quarterly KDE Layers

```{r}
battles_kde <- density(battles_ppp_com, sigma=bw.diggle, edge=TRUE, kernel="gaussian") 

plot(battles_kde)

```
