---
title: "Application of Spatial and Spatio-temporal Point Patterns Analysis to discover the geographical distribution of Armed Conflict in Myanmar"
author: "Santhya Selvan"
date: "August 30, 2024"
---

## 1.0 Overview

......

#### 1.1 Loading R packages

These are the packages that I will be using in this exercise.

**sf**: Used in spatial data wrangling

**tidyverse**: Used in data wrangling for non-spatial data

**raster**: For reading, writing and manipulating raster data

**tmap**: For functions relating to mapping point patterns

**spatstat**: Provides functions for spatial point process analysis

**sparr**: Provides functions for spatio-temporal point process analysis

```{r}
pacman::p_load(sf, raster, tidyverse, spatstat, sparr, tmap)
```

## 2.0 Data Preparation

In this section, I will be retrieving the required data and performing the necessary data wrangling methods in order to transform the data into a form that is best suitable for our analytic needs for this exercise. The datasets used in this exercise are:

-   `Myanmar_All_2021-24.csv`: A csv file downloaded from ACLED, which contains all the details of the internal conflicts in Myanmar from 2021-2024. Note that the file name has been renamed for ease of use.

-   `MBoundary`: A shapefile downloaded from Myanmar Information Management Unit, MIMU. This shapefile maps out the boundary of Myanmar. I have also renamed this file for ease of use.

#### 2.1 Importing the Data

importing the base map of myanmar/ geospatial data. transform the crs to that of Myanmar.

```{r}
boundary <- st_read(dsn="data/raw",
                    layer = "MBoundary") %>% 
  st_transform(crs=32646)

st_crs(boundary)
```

```{r}
regions_sf <- st_read(dsn="data/raw",
                      layer="regions") %>% 
  st_transform(crs = 32646)
```

importing and reading the csv file into a dataframe. transform the crs to that of Myanmars

```{r}
all_sf <- read_csv('data/raw/Myanmar_All_2021-24.csv')%>% 
  st_as_sf(coords = c('longitude', 'latitude'),
           crs=4326) %>% 
  st_transform(crs=32646)

st_crs(all_sf)
```

filter out the event types we will be focusing on in the exercise and save them into different dataframes.

```{r}
battles_sf <- all_sf %>%  filter(event_type == 'Battles')

exp_sf <- all_sf %>%  filter(event_type == "Explosions/Remote violence")

civViolence_sf <- all_sf %>%  filter(event_type == "Violence against civilians")

riots_sf <- all_sf %>%  filter(event_type=='Riots')

strat_dev_sf <- all_sf %>%  filter(event_type == 'Strategic Developments')


```

mapping out the spatial data and the spatial points by the event type to observe for any areas of interest.

```{r}
tm_shape(boundary) + tm_polygons() + tm_shape(regions_sf) + tm_polygons() + tm_shape(all_sf) + tm_dots() + tm_facets(by='event_type', free.coords = FALSE, drop.units= TRUE)
```
