---
title: "In Class Exercise 5- Geographically Weighted Summary Statistics"
author: "Santhya Selvan"
date: "September 16 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## 5.0 Getting started

Firstly, we will download the necessary packages into our environment

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, GWmodel)
```

#### 5.0.1 Downloading the data

We will then read the datafiles into sf objects

```{r}
hunan_sf<- st_read(dsn='data/geospatial',
                layer='Hunan')
```

```{r}
hunan2012_sf<- read_csv("data/aspatial/Hunan_2012.csv")
```

We will now perform a left join of the 2 sf objects based on their common identifiers

```{r}
hunan_sf<-left_join(hunan_sf, hunan2012_sf) %>% 
  select(1:3, 7, 15, 16, 31, 32)

```

It is recommended that we check for common columns with the same values before we perform the join, so that the tables can be joined using the column. We would also perform the join first before looking at the columns that are necessary for us and selecting them accordingly.
